---
title: "Codon rate screen with evolution"
author: "Alex Lukasiewicz"
date: "1/25/2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
library(gganimate)
library(viridis)
library(transformr)
```

An initial screen of codon speeds yielded the following:
```{r}
fast_slow <- read.csv("../scripts/transcript_stats_011619.csv", header = FALSE)
fast_slow <- as.data.frame(fast_slow[c(1:3)])
colnames(fast_slow) <- as.character(unlist(fast_slow[1,]))
fast_slow <- fast_slow[-1, ]
fast_slow[, 1] <- as.numeric(as.character( fast_slow[, 1] ))
fast_slow[, 2] <- as.numeric(as.character( fast_slow[, 2] ))
fast_slow[, 3] <- as.numeric(as.character( fast_slow[,3]  ))
fast_slow <- as.tibble(fast_slow)
fast_slow <- mutate(fast_slow, ratio_slow_fast = slow_rt/fast_rt)
fast_slow <- arrange(fast_slow, prod_rt)
ratio_plot <- ggplot(fast_slow, aes(x = prod_rt, y= ratio_slow_fast, group = 1)) + geom_point(aes(colour = prod_rt)) + scale_color_viridis() + ylab('ratio of slow to fast codon speed') + xlab('production rate') + theme(legend.position = 'none') +
  ggtitle('Protein production rates relative to codon speed') +
  theme_bw()
plot(ratio_plot)
```

To optimize transcripts at a variety of protein production rates, evolved each rate ratio for 50 generations. plot: final rates, overall stochasticity of rate ratio (measure changes in production rate for each ratio)
```{r}
evolved_rate_array <- read.csv("../data/codon_optimization_data/final_fitness_all.csv", header = FALSE)
evolved_rate_df <- as.data.frame(evolved_rate_array)
evolved_rate_df[, 1] <- as.numeric(as.character( evolved_rate_df[, 1] ))
evolved_rate_df[, 2] <- as.numeric(as.character( evolved_rate_df[, 2] ))
evolved_rate_df[, 3] <- as.numeric(as.character( evolved_rate_df[,3]  ))
evolved_rate_df <- mutate(evolved_rate_df, ratio_slow_fast = evolved_rate_df$V1/evolved_rate_df$V2)
evolved_rate_df <- arrange(evolved_rate_df, ratio_slow_fast)
ratio_plot <- ggplot(evolved_rate_df, aes(x = V3, y= ratio_slow_fast, group = 1)) + geom_point(aes(colour = V3)) + scale_color_viridis() + ylab('ratio of slow to fast codon speed') + xlab('production rate') + theme(legend.position = 'none') +
  ggtitle('Evolved protein production rates relative to codon speed') + 
  labs(color = 'prod_rt') +
  theme_bw()
plot(ratio_plot)
save_plot("array_evolved_production_rates.png", ratio_plot, base_aspect_ratio = 1.3)
```

To determine the overall variation in protein production as a factor of the ratio of slow:fast codon speed, the average change in protein production rate was calculated. The transcript with the minimal variation in production rates 

```{r, include = FALSE}
transcript_opt <- read.csv('../data/codon_optimization_data/codon_opt.csv', header = TRUE)
transcript_opt <- as.data.frame(transcript_opt)
transcript_summary <- transcript_opt %>% group_by(slow.codon.speed,change.in.production.rate) %>% summarise(mean = mean(change.in.production.rate), sd = sd(change.in.production.rate))
transcript_summary <- arrange(transcript_summary, mean, slow.codon.speed)
#rate_plot <- ggplot(transcript_opt, aes(y = production.rate, x = generation, group = slow.codon.speed)) + geom_line(aes(colour = slow.codon.speed))
#plot(rate_plot)
```

Additionally, this may be possible by plotting the difference between the initial production rate and the final 

```{r}
dif <- read.csv('../data/codon_optimization_data/change_in_rt.csv',header = TRUE)
dif <- as.data.frame(dif)
dif_plot <- ggplot(dif, aes(x= ï..min.speed, y = change.in.production.rate, group = 1)) + geom_point(aes(color = ï..min.speed)) + scale_color_viridis() + xlab('slow codon speed') + ylab('change in rate from generations 1 to 50') + theme(legend.position = 'none') + ggtitle('Overall changes in production rate') + labs(color = 'slow codon speed') + theme_bw()
plot(dif_plot)
#dif_plot <- dif_plot + geom_smooth(method = 'lm', se = FALSE, fullrange = TRUE, color = '#2ca25f') 
save_plot("changes_evolved_production_rates.png", dif_plot, base_aspect_ratio = 1.3)

```

Next, to observe the changes in codon usage for each rate pairing over the first 50 generations, the percentage of slow and fast codons was tracked over time

##Well, what if we evolve each rate pairing for 500 generations instead??

```{r}
transcript_stats<- read.csv("../data/presentation_datasets/transcript_stats_500_gen.csv", header = FALSE, strip.white = TRUE)
transcript_stats <- as.data.frame(transcript_stats[c(2:8)])
colnames(transcript_stats) <- as.character(unlist(transcript_stats[1,]))
transcript_stats <- transcript_stats[-1, ]
transcript_stats[, 1] <- as.numeric(as.character( transcript_stats[, 1] ))
transcript_stats[, 2] <- as.numeric(as.character( transcript_stats[, 2] ))
transcript_stats[, 3] <- as.numeric(as.character( transcript_stats[, 3] ))
transcript_stats[, 7] <- as.numeric(as.character( transcript_stats[, 7] ))
transcript_stats
transcript_stats2 <- transcript_stats %>% arrange(`Min Rate`, Generation,Pct_fast) %>% group_by(`Min Rate`) %>% mutate(codon_dif = last(Pct_fast)- Pct_fast)
transcript_stats_difference <- transcript_stats2 %>% filter(Generation == 0) %>% arrange(`Min Rate`)
difference_plot <- ggplot(transcript_stats_difference, aes(y = codon_dif, x = `Min Rate`, group = 1)) + geom_point(aes(colour = `Min Rate`)) + scale_color_viridis() + labs(title = 'Change in fast codon usage over 500 generations', x = 'slow codon rate',  y = '∆ Fast Codons in Transcript', color = 'slow codon rate' ) + geom_hline(yintercept=0, linetype="dashed", color = "black") + theme_bw() 
difference_plot
save_plot("change_in_fast_codon_usage_500_gen.png", difference_plot, base_aspect_ratio = 1.3)
#geom_smooth(method = 'lm', se = FALSE, fullrange = TRUE, color = '#2ca25f', size = 0.5) + 
```

```{r, include = FALSE}
slow_codons <- ggplot(transcript_stats, aes(x = transcript_stats[,1], y= transcript_stats[,3], group = transcript_stats[,7])) + geom_line(aes(colour = transcript_stats[,7])) + labs(title = 'changes in fast codon usage over 500 generations', x = 'generations',  y = '% fast codons', color = 'slow codon rate' ) + theme_bw() 
slow_codons 
#slow_codons <- slow_codons +  transition_states(transcript_stats[,7], state_length = 1, wrap = TRUE)
```


##Binning slow codon rates into percentiles and graphing 

```{r}
transcript_stats_bin <- transcript_stats %>% mutate(percentile = cut(`Min Rate`, 
                                       breaks = c(0,10,20,30,40,50,60,70,80,90,100), 
                                       labels = c("10","20","30","40","50","60","70","80","90","100")))
transcript_stats_average <- transcript_stats_bin %>% group_by(Generation,percentile) %>% summarize(mean = mean(Pct_fast), sd= sd(Pct_fast))
percentile_plot <- ggplot(transcript_stats_average, aes(x = Generation, y= mean, group = percentile)) + geom_line(aes(color = percentile)) + scale_color_viridis_d() + labs(title = 'changes in fast codon usage over 500 generations', x = 'generations',  y = '% fast codons in transcript', color = 'binned slow codon rate' ) + theme_bw()
percentile_plot
save_plot("change_in_fast_codon_usage_500_gen_lineplot.png", percentile_plot, base_aspect_ratio = 1.3)

```

```{r}
transcript_stats<- read.csv("../data/presentation_datasets/transcript_stats_all_new.csv", header = TRUE)
head(transcript_stats)
transcript_stats_bin <- transcript_stats %>% mutate(percentile = cut(Min, 
                                       breaks = c(0,10,20,30,40,50,60,70,80,90,100), 
                                       labels = c("10","20","30","40","50","60","70","80","90","100")))
head(transcript_stats_bin)
transcript_stats_average <- transcript_stats_bin %>% group_by(Generation,percentile) %>% summarize(mean = mean(Fast_Count), sd= sd(Fast_Count))
head(transcript_stats_average)
percentile_plot <- ggplot(transcript_stats_average, aes(x = Generation, y= mean, group = percentile)) + geom_line(aes(color = percentile)) + scale_color_viridis_d() + labs(title = 'changes in fast codon usage over 1000 generations', x = 'generations',  y = 'fast codons in transcript', color = 'binned slow codon rate' ) + theme_bw()
percentile_plot
save_plot("change_in_fast_codon_usage_1000_gen_lineplot.png", percentile_plot, base_aspect_ratio = 2.0)

```

Now to visualize the difference in fast codon usage from generation 1 to 1000

```{r}
transcript_stats2 <- transcript_stats %>% arrange(Min, Generation,Fast_Count) %>% group_by(Min) %>% mutate(codon_dif = last(Fast_Count)- Fast_Count)
head(transcript_stats2)
transcript_stats_difference <- transcript_stats2 %>% filter(Generation == 0) %>% arrange(Min)
difference_plot <- ggplot(transcript_stats_difference, aes(y = codon_dif, x = Min, group = 1)) + geom_point(aes(colour = Min)) + scale_color_viridis() + labs(title = 'Change in fast codon usage over 1000 generations', x = 'slow codon rate',  y = '∆ Fast Codons in Transcript', color = 'slow codon rate' ) + geom_hline(yintercept=0, linetype="dashed", color = "black") + theme_bw() 
difference_plot
save_plot("change_in_fast_codon_usage.png", difference_plot, base_aspect_ratio = 1.3)

```
But, what about changes in protein production??

```{r}
evolved_rate_array <- read.csv("../data/presentation_datasets/transcript_stats_all_new.csv", header = TRUE)
evolved_rate_df <- as.data.frame(evolved_rate_array)
evolved_rate_df <- arrange(evolved_rate_df, Min)
ratio_plot <- ggplot(evolved_rate_df, aes(x = Min, y= Prod_rt, group = 1)) + geom_point(aes(colour = Min)) + scale_color_viridis() + ylab('production rate') + xlab('slow codon rate') + theme(legend.position = 'none') +
  ggtitle('Evolved protein production rates relative to codon speed') + 
  labs(color = 'slow codon rate') +
  theme_bw()
plot(ratio_plot)
save_plot("array_evolved_production_rates.png", ratio_plot, base_aspect_ratio = 1.3)

```


```{r}
evolved_rate_array <- read.csv("../data/presentation_datasets/final_fitness_1000_gen.csv", header = TRUE)
evolved_rate_df <- as.data.frame(evolved_rate_array)
head(evolved_rate_array)
ratio_plot <- ggplot(evolved_rate_df, aes(x = Min, y= Rate, group = 1)) + geom_point(aes(colour = Min)) + scale_color_viridis() + ylab('protein production rate') + xlab('slow codon rate') + theme(legend.position = 'none') +
  ggtitle('Evolved protein production rates relative to codon speed') + 
  labs(color = 'slow codon rate') +
  theme_bw() + geom_smooth(method = 'loess', se = FALSE, fullrange = TRUE, color = '#2ca25f', size = 0.5)
plot(ratio_plot)
save_plot("array_evolved_production_rates.png", ratio_plot, base_aspect_ratio = 1.3)

```