---
title: "Varying Ribosome Density in Simulation"
author: "Alex Lukasiewicz"
date: "2/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
```

### Varying Ribosome copy number parameter in Pinetree
The availability of ribosomes in the translation simulation may have contributed to an increased chance of collision on the transcript, resulting in the observation that higher speed transcripts attaining a lower overall production rate. By decreasing the copy number of ribosomes in the simulation, we can observe whether higher speed ratios achieve a higher production rate. In effect we are tuning the system to remove the collision penalty. 

```{r, echo = FALSE}
evolved_rate_array <- read.csv("../data/presentation_datasets/final_fit_N10_R50.csv", header = TRUE)
evolved_rate_df <- as.data.frame(evolved_rate_array)
names(evolved_rate_df) <- c("Min","Max","Rate")
ratio_plot <- ggplot(evolved_rate_df, aes(x = Min, y= Rate, group = 1)) + geom_point(aes(colour = Min)) + scale_color_viridis() + ylab('protein production rate') + xlab('slow codon rate') + theme(legend.position = 'none') +
  ggtitle('Protein production rates: Ribosome count set to 50') + 
  labs(color = 'slow codon rate') +
  theme_bw() + geom_smooth(method = 'loess', se = FALSE, fullrange = TRUE, color = '#2ca25f', size = 0.5)
plot(ratio_plot)
save_plot("array_evolved_production_rates_R50.png", ratio_plot, base_aspect_ratio = 1.3)
```


```{r, echo = FALSE}
evolved_rate_array <- read.csv("../data/presentation_datasets/final_fit_R10.csv", header = TRUE)
evolved_rate_df <- as.data.frame(evolved_rate_array)
ratio_plot <- ggplot(evolved_rate_df, aes(x = Min, y= Rate, group = 1)) + geom_point(aes(colour = Min)) + scale_color_viridis() + ylab('protein production rate') + xlab('slow codon rate') + theme(legend.position = 'none') +
  ggtitle('Protein production rates: Ribosome count set to 10') + 
  labs(color = 'slow codon rate') +
  theme_bw() + geom_smooth(method = 'loess', se = FALSE, fullrange = TRUE, color = '#2ca25f', size = 0.5)
plot(ratio_plot)
save_plot("array_evolved_production_rates_R10.png", ratio_plot, base_aspect_ratio = 1.3)
```


```{r, echo = FALSE}
evolved_rate_array <- read.csv("../data/presentation_datasets/final_fit_R5.csv", header = TRUE)
evolved_rate_df <- as.data.frame(evolved_rate_array)
names(evolved_rate_df) <- c("Min","Max","Rate")
ratio_plot <- ggplot(evolved_rate_df, aes(x = Min, y= Rate, group = 1)) + geom_point(aes(colour = Min)) + scale_color_viridis() + ylab('protein production rate') + xlab('slow codon rate') + theme(legend.position = 'none') + 
  ggtitle('Protein production rates: Ribosome count set to 5') + 
  labs(color = 'slow codon rate') +
  theme_bw() + geom_smooth(method = 'loess', se = FALSE, fullrange = TRUE, color = '#2ca25f', size = 0.5)
plot(ratio_plot)
save_plot("array_evolved_production_rates_R5.png", ratio_plot, base_aspect_ratio = 1.3)
```

The relationship between production rate and speed, without the influence of ribosome collisions appears to be more exponential than linear. 

